\documentclass{article}

\usepackage{../mathclub}

\title{Number Theory - Week 8}
\author{}
\date{November 14, 2022}

\begin{document}

\section{Introduction}

    Welcome back to Math Club in week 8! 
    Today we will be playing with some concepts in elementary number theory. 
    We will cover modular arithmetic as well as some stuff about primes and divisibility, all of which will be helpful in not just math competitions but \textit{every single one of your math classes moving forward}. 
    We have provided some useful definitions and theorems for your convenience.

    \begin{definition}
        Let \(x\), \(y\), and \(n\) be integers. 
        We say that \(x\) is congruent (or equivalent) to \(y\) modulo (or mod) \(n\) if \(x\) and \(y\) differ by a multiple of \(n\). 
        We write \(x\equiv y\pmod{n}\).
    \end{definition}

    \begin{definition}
        Let \(a,b\in\ZZ\). 
        We say that $b$ divides $a$, in symbols $b \mid a$, if there exists an integer $k$ such that $a=kb$. We write $b\nmid a$ to mean ``$b$ does not divide $a$.''
    \end{definition}

    \begin{definition}
        Let \(a,b\in\NN\). 
        Then their greatest common divisor, written \((a,b)\) or \(\gcd(a,b)\), is the largest integer that divides both \(a\) and \(b\). 
        Their least common multiple, written \(\lcm(a,b)\), is the smallest positive integer that is a multiple of both \(a\) and \(b\).

        If \(\gcd(a,b)=1\), we say that \(a\) and \(b\) are coprime or relatively prime.
    \end{definition}
    
    \begin{theorem}
        The following hold:
        \begin{enumerate}
            \item If \(g\mid a\) and \(g\mid b\), then \(g\mid ax+by\) for any integers \(x,y\)
            \item \(\gcd(a+b,b)=\gcd(a,b)\)
            \item \(\gcd(a,b)\cdot\lcm(a,b)=a\cdot b\)
            \item If \(p\) is prime and \(p\mid ab\), then \(p\mid a\) or \(p\mid b\)
        \end{enumerate}
    \end{theorem}
    
    \begin{theorem}[Division Algorithm]
    Let $a, b\in \ZZ$ with $b\neq 0$. There exist unique integers $q,r$ such that $a=bq+r$ and $0 \leq r < |b|$.
    \end{theorem}

    Here are some useful results that will help you in your problem-solving endeavors.

    \begin{theorem}[Fermat's Little Theorem]
        Let \(a\in\ZZ\) and let \(p\) be a prime number. 
        Then
        \[a^p\equiv a\pmod{p}.\]
    \end{theorem}

    \begin{theorem}[Bezout's Identity/Lemma]
        Let \(a,b\in\NN\). 
        Then there exist \(x,y\in\ZZ\) such that
        \[ax+by=\gcd(a,b).\]
        and moreover, $\gcd(a,b)$ is the least positive integer expressible in this form.
    \end{theorem}
    
    \begin{theorem}[Fundamental Theorem of Arithmetic]
    Every positive integer $n>2$ can be represented in exactly one way as the product of prime powers
    \[n=p_1^{k_1}p_2^{k_2}\cdots p_r^{k_r}=\prod_{i=1}^r p_i^{k_i}\] where $p_1<p_2<\cdots<p_r$ are primes and the $k_i$ are positive integers.
    \end{theorem}

\section{Problems}

    Try to prove some of the results given in the introduction section.

    \begin{exercise}
        Prove that there are infinitely many prime numbers.
    \end{exercise}

    \begin{exercise}
        Prove that there are infinitely many primes \(p\) such that \(p\equiv 3\pmod{4}\).
    \end{exercise}

    \begin{exercise}
        Let \(n\) be an odd positive integer not divisible by 3. Show that \(n^2-1\) is divisible by 24.
    \end{exercise}

    \begin{exercise}
        A \textit{palindrome} is a positive integer that reads the same forward and backward, like 2552 or 1991. 
        Find a positive integer greater than 1 that divides all four-digit palindromes.
    \end{exercise}

    \begin{exercise}
        Find and integer \(c\) such that \(x^2+18x+c\) is a perfect square for all integers \(x\). 
        Prove that this choice of \(c\) is unique. 
    \end{exercise}

    \begin{exercise}[Due to Sun-tzu, 3rd century]
        There are certain things whose number is unknown. 
        If we count them by threes, we have two left over; by fives, we have three left over; and by sevens, two are left over. 
        How many things are there?
    \end{exercise}

    \begin{exercise}
        Show that any two consecutive Fibonacci numbers are relatively prime. 
        Recall that the Fibonacci numbers are defined recursively by \(F_1=F_2=1\) and \(F_n=F_{n-1}+F_{n-2}\) for \(n\geq 3\).
    \end{exercise}

    \begin{exercise}
        Determine whether there exist three positive integers \(a,b,c\) such that \(a+b\), \(b+c\), and \(a+c\) are all pairwise distinct prime numbers.
    \end{exercise}

    \begin{exercise}
        Let \(k=2020^2+2^{2020}\).
        What is the last digit of
        \[2^k+k^2\mathord{?}\]
    \end{exercise}

    \begin{exercise}
        Show that if \(a^2+b^2=c^2\), then \(3\mid ab\).
    \end{exercise}

    \begin{exercise}
        Show that the fraction \(\dfrac{21n+4}{14n+3}\) is irreducible for all positive integers \(n\).
    \end{exercise}

    \begin{exercise}
        Prove that the number \(n=1,280,000,401\) is composite.
    \end{exercise}

    \begin{exercise}
        Let \(N\) be a number with nine distinct non-zero digits, such that, for each \(k\) from 1 to 9 inclusive, the first \(k\) digits of \(N\) form a number that is divisible by \(k\). 
        Find \(N\).
    \end{exercise}

    \begin{exercise}[2000 B2]
        Prove that the expression
        \[\frac{\gcd(m,n)}{n}\binom{n}{m}\]
        is an integer for all pairs of integers \(n\geq m\geq 1\).
    \end{exercise}

    \begin{exercise}[2017 A1]
        Let \(S\) be the smallest set of positive integers such that
        \begin{enumerate}
            \item[(a)] \(2\in S\),
            \item[(b)] \(n\in S\) whenever \(n^2\in S\),
            \item[(c)] \((n+5)^2\in S\) whenever \(n\in S\).
        \end{enumerate}
        Which positive integers are not in \(S\)?
    \end{exercise}
    
    \begin{exercise}
        If $a\in \NN$ and $p$ is a prime number for which $p$ divides $(a^7-1)/(a-1)$, prove that either $p \equiv 1 \pmod{7}$ or $p=7$.
    \end{exercise}
    
    \begin{exercise}
        Find all integer solutions of the equation
        \[\frac{a^7-1}{a-1} = b^5-1\]
    \end{exercise}
    
    \begin{exercise}
        If $a\equiv b \pmod{n}$ prove that $a^n \equiv b^n \pmod{n^2}$.
    \end{exercise}

    \begin{exercise}[2011 A4]
        For which positive integers n is there an \(n\times n\) matrix with integer entries such that every dot product of a row with itself is even, while every dot product of two different rows is odd?
    \end{exercise}
    
    \begin{exercise}
        Let $p$ be a prime. Show that there are infinitely many positive integers $n$ such that $p$ divides $2^n-n$.
    \end{exercise}
    
    \begin{exercise}
        Let $n>1$ be a positive integer. Prove that 
        \[1+\frac{1}{2}+\frac{1}{3}+\cdots + \frac{1}{n}\] is not an integer.
    \end{exercise}

    \begin{exercise}[2001 A5]
        Prove that there are unique positive integers \(a\) and \(n\) such that 
        \[a^{n+1}-(a-1)^n = 2001.\]
    \end{exercise}

\end{document}